---
title: "Masters Homework 2"
author: "Michelle Murata"
date: "7/21/2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


>- Find summary information about the total blood mercury levels in our NYC HANES data (hint: names(nyc) will call the names of each column). What is the max value, what is the minimum value.  

```{r}
nyc <- read.csv("https://raw.githubusercontent.com/fredlapolla/RVilcekMaster23/master/NYC_HANES_DIAB.csv")
print(names(nyc))
summary(nyc$MERCURYTOTALBLOOD)
# max value is 129.2344 and min value is 0.0919
```

>- Bonus/Challenge: What is the Standard Deviation and skewness of the blood mercury?

```{r}
sd(nyc$MERCURYTOTALBLOOD)
# unsure if there's a command for skew in base R
```



>- Install a package and load, such as "ggplot2" or "dplyr"

```{r}
install.packages("dplyr")
library(dplyr)
```





>- What two commands are needed to install and use a package in R?

```{r}
# install.packages() and library()
```




>- When you read in a .csv file into R, what class is the object that you read in (like nyc)?
The object is a dataframe


>- What command can you use to determine what type of data is in a column?

```{r}
# class(dataframe$column)
```


>- Provide a command to find the type of data in the DX_DBTS column of nyc. 

```{r}
class(nyc$DX_DBTS)
```


>- Provide the same command the type of data but use the other way of indexing from what you did above. 

```{r}
class(nyc[["DX_DBTS"]])
```



>- What would happen if we had chosen the first 4 columns of NYC HANES and made a matrix? What would the class of the cells in this matrix be? Why?

```{r}
# They would be of the character class because R requires all classes in a matrix to be the same, and character is the class that can accomodate the first 4 columns
```


>- If we make a vector,vec<- c(3,9,4,7,8,9), and multiply it by WannaMatrix, what is the value in the third row, third column? 

```{r}
WannaMatrix <- nyc[5:10, 6:9]
WannaMatrix <- as.matrix(WannaMatrix,nrow = 6, ncol = 4)
print(WannaMatrix)
vec<- c(3,9,4,7,8,9)
product <- vec %*% WannaMatrix
print(product[1,3])
```



>- What happens if instead we made a vector vec2 <- c(3,9,4,7,8) and multiply it by WannaMatrix, what is the third row, third column value? What is going on? 

```{r}
print(WannaMatrix)
vec2 <- c(3,9,4,7,8)
product2 <- vec2 %*% WannaMatrix
print(product2[1,3])
# Returns an error because the dimensions don't work for matrix multiplication (the number of rows in the matrix has to equal the length of the vector)
```


>- Look up the arguments for read.csv

```{r}
help(read.csv)
```


>- Use apply() to find the median value of CADMIUM, LEAD CHOLESTEROLTOTAL, GLUCOSESI, CREATININESI, and CREATININE

```{r}
print(apply(nyc[,12:19],2,median,na.rm=TRUE))
```


>- Do the same using sapply()

```{r}
print(sapply(nyc[,12:19],median,na.rm=TRUE))
```

>- Use a for loop to find the median value of columns 16:19

```{r}
calc_median <- numeric(4)
print(calc_median)
for (i in 16:19) {
  calc_median[i-15] <- median(nyc[[i]], na.rm=TRUE)
}
print(calc_median)
```


>- Make a for loop to find the standard deviation (hint sd() ) of CHOLESTEROLTOTAL, GLUCOSESI, CREATININESI, AND CREATININE

```{r}
calc_sd <- numeric(4)
print(calc_sd)
for (i in 16:19) {
  calc_sd[i-15] <- sd(nyc[[i]], na.rm=TRUE)
}
print(calc_sd)
```


>- Bonus/Challenge Create a vector in which to put the standard deviations 

```{r}
# see above
```


>- Use ifelse to create a new variable of high and normal A1C dichotomizing with 7 as a threshold for high levels

```{r}
A1C_category <- ifelse(nyc$A1C>7, "high", "normal")
print(A1C_category)
```
###################################################################################OPTIONAL##############################################################################################

>- Subset to only have females (gender == "female")

```{r}

```


>- Create a subset of males and find the median total cholesterol level

```{r}

```


>- Try to create a summary of the mean A1C using two factors, GENDER and DX_DBTS

```{r}

```

Find the class of the first three columns of data. What are they?

```{r}

```

